---
title: "ARIMA Models"
output: html_notebook
---

Plot white noise

```{r}
n <- 100
time <- 1:n
set.seed(2018)
W <- rnorm(n, mean=0, sd = 0.2)
plot(time, W, type="l", ylim=c(-1,1))
```

Estimate autocorr. Blue dashed lines are pointwise confidence intervals for autocorr, NOT dependent on h

```{r}
rho <- acf(W)
rho$acf
```

AR(1) process

```{r}
phi <- 0.9; sigma <- 0.2
set.seed(2018)
W <-rnorm(n, mean=0, sd=sigma)
Y <- rep(0, n)
Y[1] <- W[1] # initialise
for (t in 2:n) Y[t] <- phi * Y[t-1] + W[t]
plot(time, Y, type="l", col="blue", ylim=c(-1,1))
points(time, W, col="orange", type="l")
legend("topleft", c("Y", "W"), col=c("blue","orange"), lty=1)
acf(Y)
```

```{r}
plot(arima.sim(list(order=c(1,0,0), ar=0.9), n=100, sd=0.2), ylab="Y")
plot(arima.sim(list(order=c(1,0,0), ar=-0.9), n=100, sd=0.2), ylab="Y")
```

```{r}
theta <- 0.9; sigma <- 0.2
set.seed(2018)
W <-rnorm(n, mean=0, sd=sigma)
Y <- rep(0, n)
Y[1] <- W[1]/(sqrt(1-phi^2)) # initialise at random, divide through for stationarity
for (t in 2:n) Y[t] <- theta * W[t-1] + W[t]
plot(time, Y, type="l", col="blue", ylim=c(-1,1))
points(time, W, col="orange", type="l")
legend("topleft", c("Y", "W"), col=c("blue","orange"), lty=1)
acf(Y)
```

Use arima for MA

```{r}
acf(arima.sim(list(order=c(0,0,1), ma=0.9), n=100, sd=0.2))
acf(arima.sim(list(order=c(0,0,1), ma=0.2), n=100, sd=0.2))
```

Now combine AR and MA into ARIMA

```{r}
set.seed(2018)
Y <- arima.sim(list(order=c(2,0,1), ar=c(0.9, -0.3), ma=0.5), n=100, sd=0.2)
plot(Y, type="l")
acf(Y)
```

ARIMA Model with d=1 

```{r}
set.seed(2018)
Y <- arima.sim(list(order=c(2,1,1), ar=c(0.9, -0.3), ma=0.5), n=100, sd=0.2)
plot(Y, type="l")
```

Can revocer ARMA by taking difference (values are lagged by 1)

```{r}
Y <- diff(Y)
plot(Y, type="l")
acf(Y)
```

```{r}
filePath <- "https://raw.githubusercontent.com/AJCoca/SLP19/master/"
fileName <- "wait.txt"
dat <- read.table(paste0(filePath, fileName), header=T)
wait <- dat$wait
```

Plot time series

```{r}
time <- 1:length(wait)
plot(time, wait, type="l")
rho <- acf(wait)
rho
```

Lags go up to 5 --> not an MA(1)

```{r}
wait.ts <- ts(wait)
wait.ar <- arima(wait.ts, order=c(1,0,0))
wait.ar
```

```{r}
plot(wait.ar$residuals)
acf(wait.ar$residuals)
```

Now fit MA

```{r}
wait.ma <- arima(wait.ts, order=c(0,0,5))
wait.ma
plot(wait.ma$residuals,type="l")
acf(wait.ma$residuals)
```

```{r}
filePath <- "https://raw.githubusercontent.com/AJCoca/SLP19/master/"
fileName <- "calls.txt"
calls_number <- read.table(paste0(filePath, fileName), header=T)
head(calls_number)
```

```{r}
?ts
```


```{r}
calls_series<-ts(data=calls_number,start=c(7,1),frequency=4)
plot(calls_series,xlab="hours")
acf(calls_series)
```

```{r}
install.packages("tseries")
library(tseries)
```

Tests H0: stationary vs H1: non-stationary

```{r}
kpss.test(calls_series)
```

Use pacf to infer order of AR(p) model (so in this case --> suggests order at most 2)

```{r}
pacf(calls_series)
```

```{r}
for (p in 0:2) for (q in 0:2){
 cat("p = ", p, "q = ", q, "AIC = ",
 AIC(arima(calls_series, order=c(p,0,q), include.mean=F)), "\n")
} 
```

```{r}
call.ts <- arima(calls_series, order=c(2,0,2), include.mean=FALSE)
call.ts
```

```{r}
install.packages("forecast")
library(forecast)
```

Use auto arima to choose best ARIMA model based on AIC

```{r}
auto.arima(calls_series, ic="aic", seasonal = FALSE, trace=TRUE)
```

ARIMA gives Inf for ARIMA(2,0,2) as abs value of roots of AR & MA equations are close to 1.

```{r}
abs(polyroot(c(1,-coef(call.ts)[c("ar1", "ar2")])))
abs(polyroot(c(1,coef(call.ts)[c("ma1", "ma2")])))
```
```{r}
calls.arma <- arima(calls_series, order=c(1,0,1), include.mean=F)
fc <- forecast(calls.arma, 48, level=95)
plot(fc, shaded=F)
```
```{r}
?forecast 
```

Stock prices

```{r}
library(quantmod)
getSymbols("SLP", from="2010-01-01", to="2017-01-01")
slp = SLP[,6]
head(slp)
plot(slp)
```

```{r}
plot(log(slp))
```

Check whether log of series is stationary.

```{r}
diff_slp <- diff(log(slp)) 
log_slp <- log(slp)
kpss.test(log_slp)
kpss.test(diff_slp)
```

```{r}
plot(log_slp)
plot(diff_slp)
acf(diff_slp[-1])
pacf(diff_slp[-1])
```

Use auto arima to select best ARIMA model.

```{r}
diff_ts <- ts(diff_slp)
log_ts <- ts(log_slp)
auto.arima(diff_ts, ic="aic")
auto.arima(log_ts, ic="aic")
```

Want to take one step ahead forecasting and compare to actual return - ie backtest the model

```{r}
getSymbols("SLP", from="2010-01-01", to="2020-03-11")
slp2 = SLP[,6]
head(slp2)
```

```{r}
diff_slp2 <- diff(log(slp2)) 
log_slp2 <- log(slp2)
kpss.test(log_slp2)
kpss.test(diff_slp2)
```

```{r}
X <- as.numeric(diff_slp2)
n <- length(slp2)
pred <- pred_low <- pred_high <- rep(0,n)
test_start = 1764 # first day of 2017, as there are 250 business days
for (t in test_start:n){
# we use Arima(), which is a wrapper of arima() but allows us to include drift
model <- Arima(X[1:(t-1)], order=c(0,1,1), include.drift=TRUE)
tmp <- forecast(model, 1, level=95)
pred[t] <- tmp$mean; pred_low[t] <- tmp$lower; pred_high[t] <- tmp$upper
}
```

```{r}
backtest <- data.frame(actual = X[test_start:n], pred = pred[test_start:n])
backtest.lm <- lm(actual~pred, data=backtest)
with(backtest, plot(pred, actual))
summary(backtest.lm)
abline(backtest.lm)
````



